

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(BSDA)
source("Functions.R")
knitr::opts_chunk$set(echo = TRUE)
SHOW_SOLUTIONS = TRUE
```
1.Read the ice cream, birthweight, and cholesterol data sets.

```{r, eval=TRUE}
IC.df <- read.csv("E:/DATA SCIENCE JOBS/R-Analysis Significance Test/Background/IceCream.csv", header=TRUE, as.is=TRUE)
IC.df$Sex <- as.factor(IC.df$Sex)
IC.df$Flavor <- as.factor(IC.df$Flavor)
View(IC.df)
#
BW.df <- read.csv("E:/DATA SCIENCE JOBS/R-Analysis Significance Test/Background/BirthWeight.csv", header=TRUE, as.is=TRUE)
BW.df$Smoker <- as.factor(BW.df$Smoker)
BW.df$BirthWt <- as.factor(BW.df$BirthWt)
BW.df$MAgeGT35 <- as.factor(BW.df$MAgeGT35)
View(BW.df)
#
C.df <- read.csv("E:/DATA SCIENCE JOBS/R-Analysis Significance Test/Background/Cholesterol.csv", header=TRUE, as.is=TRUE)
View(C.df)
C.df$Cereal <- as.factor(C.df$Cereal)
```
1. Find and interpret the 92% confidence interval for the population mean puzzle score.






2. Hypothesis test for the difference between mean puzzle scores by ice cream type. a. Create two subsets of puzzle scores, one for students that favor strawberry ice cream and one for students that favor chocolate ice cream. b. Test the claim that students with a preference for strawberry ice cream have higher puzzle scores than students that prefer chocolate ice cream. Use strawberry minus chocolate and a 1% significance level. Assume the population variances are not equal. c. Do you think the statistical test results from part (b) have practical significance? 3. Using the oneprop.CI function in the R4 Tutorial, find the 98% confidence interval for the proportion of smokers in the birthweight data set. Assume large sample conditions are met.






4. Consider birthweights for mothers that are smokers and nonsmokers.   a. Create a function called twoprop.HT that uses a two-sample z test to test for the difference between two population proportions assuming independent samples. b. Test the claim that the proportion of low birthweight babies is higher for mothers that smoked (use smoked – did not smoke). Use a 5% significance level. Assume large sample conditions are met. c. Are the large sample conditions met for the test in (b)? Note, you can just look at the table values instead of doing any calculations. Do you think the hypothesis test results in
(b) are valid? 
 
5. This problem was modified from here. A cross-over trial experiment was used to investigate whether eating oat bran lowered serum cholesterol levels. Twelve individuals were randomly assigned a diet that included either oat bran or corn flakes. After two weeks on the initial diet, serum cholesterol (mmol/L) was measured and then participants were “crossed-over” to the other diet. After two-weeks on the second diet, cholesterol levels were once again recorded.  a. Using a 5% significance level, test the claim that a diet that includes oat bran decreases serum cholesterol. [Use Cornflk – OatBran] b. Construct an appropriate confidence interval that could be used as an equivalent to the test in part (a). [Use Cornflake – OatBran]. Explain your choice and interpret the interval. 


 6. In clinical experiments involving different groups of independent samples, it is important that the groups be similar in the important ways that affect the experiment. In an experiment designed to test the effectiveness of paroxetine for treating bipolar depression, subjects were measured using the Hamilton depression scale with the results given below (based on data from a “Double-Blind, Placebo-Controlled Comparison of Imipramine and Paroxetine in the Treatment of Bipolar Depression,” by Nemeroff et al., American Journal of Psychiatry, Vol. 158, No. 6). Use a 0.05 significance level to test the claim that the treatment and placebo groups come from populations with the same mean. Assume equal population variances. [Use Treatment – Placebo]  
